### 2018-02-16 article/ms.tex
cant spell nobel prize without r-process enhancement

\documentclass[a4paper,fleqn,usenatbib]{mnras}

\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

\usepackage{numprint}
\usepackage{booktabs}
\usepackage{float}

\usepackage{amsbsy}
\usepackage{graphicx}	% Including figure files
\usepackage{color}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

\newcommand{\todo}[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.

\title[Discovery and origin of $r$-process stars]{On the discovery and origin of $r$-process stars in the Milky Way}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Matthew T. Miles et al.]{Matthew T. Miles,$^{1}$\thanks{E-mail: mtmil3@student.monash.edu}
	Andrew R. Casey,$^{1,2}$
	Brodie J. Norfolk,$^{1}$
	Alex J. Kemp,$^{1}$\newauthor
	John C. Lattanzio,$^{1}$
	Kevin C. Schlaufman,$^{3}$
	Alexander P. Ji,$^{4}$
	Anna Y. Q. Ho$^{5}$
	\\
	% List of institutions
	$^{1}$School of Physics and Astronomy, Monash University, Clayton Campus, Victoria 3800, Australia\\
	$^{2}$Faculty of Information Technology, Monash University, Clayton Campus, Victoria 3800, Australia\\
	$^{3}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N Charles St., Baltimore, MD 21218, USA\\
	$^{4}$Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA\\
	$^{5}$Cahill Center for Astrophysics, California Institute of Technology, MC 249-17, 1200 E California Blvd, Pasadena, CA, 91125, USA
}

\date{Accepted XXX. Received YYY; in original form ZZZ}

\pubyear{2018}

\begin{document}
	\label{firstpage}
	\pagerange{\pageref{firstpage}--\pageref{lastpage}}
	\maketitle

	\begin{abstract}
		The production of heavy elements occurs through the slow and rapid neutron capture process. Numerous sites satisfy the conditions needed for the r-process but the relative frequency of these sites is unknown, largely due to the lack of a large number of stars with strong r-process signatures. Here we report the discovery of 47 r-process enhanced stars discovered in the second LAMOST data release (DR2), more than doubling the known number in our Galaxy. After taking expected rates of core-collapse supernovae and the average r-process material ejected in each event, we find the core-collapse supernovae alone cannot explain the number of r-process stars discovered in this work. Current rate estimates of neutron-star mergers imply that should be about $9896$ r-process enhanced stars in the Milky Way, and about $0.087$ in LAMOST DR2. We find that neutron star mergers are also able to create all currently known r-process enhanced stars within a reasonable timescale. Additionally these 47 stars show no correlation to poor metallicity, providing evidence against the relationship of poor metallicity environments and the action of the r-process.
		To our knowledge, this is the first work that links the rates of neutron-star mergers to the number of r-process stars in the Milky Way.
	\end{abstract}

	\begin{keywords}
		nucleosynthesis -- r-process -- neutron star mergers -- core-collapse supernovae
	\end{keywords}

	\section{Introduction}

	Heavy elements ($Z > 30$) are synthesised through the slow (s) and rapid (r) neutron capture processes \citep[e.g.,][]{Burbidge1957,Sneden2008}.
	Rapid neutron capture occurs when the neutron density is high enough that the rate of neutron capture will occur faster than the rate at which the isotope undergoes a $\beta$-decay. Extremely high neutron densities ($>10^{24}\ \rm{free\ neutrons\ per\ cm}^{3}$) \citep[e.g.,][]{Burbidge1957,Sneden2008} are necessary for the r-process to occur. Sites meeting this criteria include core-collapse supernovae, neutron star mergers, or magnetorotationally driven supernovae. Core-collapse supernovae are relatively common throughout the Galaxy, but they only produce small amounts of r-process material. Neutron star mergers or magnetorotationally driven supernovae, however, occur much less frequently than core-collapse supernovae, but produce far more r-process material. Events such as these are theorised to account for the case of the highly r-process enhanced stars in the ultra-faint dwarf (UFD) galaxy Reticulum II \citep{Ji2016}.
	Magnetorotationally driven supernovae may not produce as much r-process material as neutron star mergers (although still more than core-collapse supernovae), they are thought to occur more frequently than neutron-star mergers \citep[e.g.,][]{Li2011, LIGO2016}, and the predicted abundance patterns make them observationally indistinguishable from neutron star mergers \citep{Ji2016}.

	Europium ($Z=63$) is entirely produced through the r-process. This makes it a very useful signature for identifying stars that show evidence of excess pollution by r-process material. Using Europium, Reticulum II found the particularly interesting result that the UFD's level of r-process enrichment was too high to have occurred through a core-collapse supernovae event.
	Historically, UFD galaxies have demonstrated some of the lowest abundance ratios of r-process elements seen in the Milky Way or its satellite systems. However, the analysis of several of Reticulum II's stars show they are greatly enriched, 2-3 orders of magnitude higher than stars found in any other UFD galaxy \citep{Ji2016}. This degree of enrichment supports the theory that a single rare event must have occurred early on in the formation history of Reticulum II. This is supported by the observed heavy element yields, which are found to be some 1000 times higher than what is achievable on average from core-collapse supernovae ejecta. It is highly improbable that 1000 supernovae could have contributed to Reticulum II, suggesting that the r-process material must be produced by a rarer astrophysical event such as a binary neutron star merger.

	While neutron star mergers or magnetorotationally driven supernovae are likely to produce significant amounts of r-process material, it does not suggest that all r-process material stems from these sources either. Rather, the rarity of these events might imply that they are also responsible for the rare numbers of stars  observed to have significant enhancements in r-process material. One could ask if neutron-star mergers could be responsible for all r-process stars known in the Milky Way? On the other hand, if we assume very little gas mixing, could core-collapse supernovae create enough material to explain a single r-process star? If neither of these prove to be true then we can confirm that there must be multiple sties to the r-process, providing a step forward in understanding the relative fraction of these mechanisms in the Milky Way.

	In order to answer these questions it is necessary to identify a large sample of r-process stars.
	However, r-process stars are very rare, which makes finding this stars difficult. For this reason we must look to massive data sets, such as the LAMOST sky survey. In this letter we report the discovery of 47 r-process enhanced stars identified from LAMOST, significantly building upon the previously known Milky Way sample of 14 \citep[e.g.,][]{masho2011,Honda2004,Hayek2009,Sneden2003,Ivans2006,Hill2002,Christlieb2004,Wako2010,placo2017}.
	In Section 2 we outline the methods used to identify r-process enhanced candidates, and our analysis methods. In Section 3 we discuss what our results imply about the sites of r-process. We provide concluding remarks in Section 4.

	\begin{figure}
		\includegraphics[width=\columnwidth]{423451}
		\caption{The spectral region around two Europium absorption lines at 4129 and 4205 Angstroms. This particular star's spectra (black) shows clear absorption at these wavelengths, as compared to the data-driven model (red).}
		\label{fig:starindex_423451}
	\end{figure}

	\begin{figure}
		\includegraphics[width=\columnwidth]{metalhistpython}
		\caption{The metallicities of the 47 candidate stars.}
		\label{fig:metallicity}
	\end{figure}

	\section{Methods}

	\subsection{Observations and Candidate Selection}

	The LAMOST survey is a low resolution optical spectroscopic survey ($R\approx1800$; 3650-9000\,\AA). The second data release from LAMOST \citep{lamost} contains about 2.2 million stellar spectra in the northern sky. Of these 2.2 million, 454,180 were identified as giant stars through a data-driven analysis known as \textit{The Cannon}. An overlap sample of 9,952 giant stars common to LAMOST and the higher-resolution ($R\approx22500$) APOGEE survey \citep{apogee} were used as a training set.
	The stellar labels used by \textit{The Cannon} included $T_{\rm{eff}}$, $\log{g}$, $\rm{[Fe/H]}$, and $\rm{[\alpha/Fe]}$. For reference, typical uncertainties for these labels are about $70\rm{K}$ for $T_{\rm{eff}}$, $0.1$ in $\rm{log_{10}{g}}$, 0.1\,dex in $\rm{[Fe/H]}$, and 0.04\,dex in $\rm{[\alpha/Fe]}$.

	We identified candidate r-process enhanced stars by searching for deviations surrounding between the pseudo-continuum-normalised flux and the best-fitting data-driven model spectrum. Specifically, we looked for deviations at the 4129\,\AA\ and 4205\,\AA\ europium absorption lines. At each transition we fitted a Gaussian profile to the flux residuals for all 454,180 giant stars. We recorded the amplitude, wavelength, and width of each profile, as well as the corresponding uncertainties on these quantities. 
	Candidate r-process stars were identified as stars that met the following criteria:

	\begin{itemize}
		\item The amplitude of the gaussian profile must be less than $\rm{A}<-0.05$;
		\item The signal to noise (S/N) must be greater than 30;
		\item The deviation from the required wavelength in question must be less than 2 \AA;
		\item The  $\chi_{\rm{r}}^{2}$ value must be less than 3, $\chi_{\rm{r}}^{2}<3$.
		\item All of these conditions must be met for both 4129 \AA\ and 4205 \AA\ for us to consider them sufficiently Eu enhanced.
	\end{itemize}

	These filters produced an initial sample of 62 candidates from the total sample of 454,180. An example candidate is shown in Figure \ref{fig:starindex_423451}. We visually scrutinised each candidate to remove false positives. After this process we were left with 47 likely r-process enhanced candidates, 22 of which stand out as very strong candidates.

	\section{Identifying the r-process}

	\subsection{Selection effects}
	The LAMOST giants we analyse belong to a broad range of metallicity ($-1.6 < \rm{[Fe/H]} < 0.6$), and don't cluster in any part of the sky, suggesting they do not belong to a single stellar system.

	The previous sample of r-process stars, before the release of this letter, almost exclusively consists of metal-poor stars \citet{Barklem2005}. This fact is not to be confused with saying that most r-process stars found should be metal-poor stars, rather that at low metallicities we know that the s-process will not contribute, so neutron capture elements formed from these stars would be only from the r-process. Over half the current known r-process stars are members of the ultra faint dwarf galaxy, Reticulum II. However, Europium we know to be totally made from the r-process, therefore even at metallicities where the s-process contributes, Europium is a valuable indicator of r-process synthesis.

	\subsection{Multiple r-process sites?}
	While it is possible that all currently known r-process enhancements can be made by neutron star mergers (we show this in Section \ref{rates}), we still observe stellar populations in the galaxy which show particularly weak enrichment from r-process elements ([Eu/Fe] < 0.7). These enhancements imply either that there is more than $1$ active r-process site, or that there is present very heterogeneous levels of mixing, such that neutron-star mergers are able to enhance stars with [Eu/Fe] abundances that range many orders of magnitude.

	\subsection{Core-collapse supernovae (CCSN)}
	Core-collapse supernova are one of the sites initially proposed to have a high enough neutron density to create elements synthesised by the r-process \citep{Burbidge1957}.

	As of yet our knowledge is incomplete on the exact mechanism that allows material with a high enough neutron density to escape the supernovae event.
	One possible mechanism is the action of high-entropy neutrino winds being released from a proto-neutron star during core-collapse. In this stage of the collapse, this neutrino wind is driving mass loss and could be instrumental in the expulsion of the r-process material in supernovae conditions. However, how much enhanced material is released from the event is very dependant on the amount of material that is allowed to fall back onto the remnant at the arrival of the reverse shock, and so often the supernovae may not release a considerable amount of material \citep{Woosley1992, Burrows1995}.

	Regardless of the exact mechanism, core-collapse supernovae seem to release $M_{\rm{Eu}}\approx10^{-7.5} M_{\odot}$ per event \citep{Argast2004} . While this may not be a particularly significant amount, core-collapse supernovae occur frequently \citep[$44700\ \rm{Gpc}^{-3} \rm{yr^{-1}}$;][]{Li2011}, and so it isn't unreasonable to suggest that, with very little mixing occurring, that much if not all r-process material could have been formed by these and recycled in the stars we present in this letter.

	\subsection{Neutron star mergers}
	\label{NSmerg}
	Neutron star mergers are another possible site of r-process element nucleosynthesis \citep{Kasen2017,Hotok2013,Drout2017}. They satisfy the conditions necessary to create a high neutron density by means of interacting tidal forces between the stars, neutron dense material is accreted off the stars, allowing r-process element synthesis to occur. This specific process produces considerably more enriched material than core-collapse supernovae ($M_{\rm{Eu}}\approx10^{-4.5} M_{\odot}$) \citep{Goriely2011}. However, the direct interaction and merger of a binary neutron star pair occurs very infrequently, with currently only one direct observation, and with a frequency estimate of $1000\ \rm{Gpc^{-3}yr^{-1}}$ \citep{LIGO2016}. The significant amount of material that these mergers release can compensate for their low frequency and so, synonymous with core-collapse supernovae, it is not unreasonable to suggest that they are responsible for most if not all known r-process material.

	The recent discovery of the UFD Reticulum II \citep{Ji2016} is evidence against the case of a single r-process site. Of nine stars observed in Reticulum II, 7 of them were shown to be highly enriched in r-process elements ($\rm{[Eu/Fe]\approx1.7}$) which suggests a single prolific event having occurred in close proximity to the galaxy early in it's formation, such as a neutron star merger. However, two of the stars observed were only weakly enhanced in r-process elements which suggests that there must have been a second, less efficient, form of r-process enhancement present.

	\subsection{The rates of r-process events and what they imply}
	\label{rates}
	From the recent advanced LIGO data \citep{LIGO2016}, we are able to find an estimate for a plausible frequency at which neutron star mergers occur, as mentioned in section \ref{NSmerg}. As well as a reliable estimate of core-collapse supernovae rates from \citet{Li2011}.

	From these two rates, we can for the first time to our knowledge, estimate the amount of r-process giants that should be present in the Milky Way, enhanced both by purely neutron star mergers, and purely core-collapse supernovae. As well as an estimate of how many enhanced stars, giants or otherwise, that should be present in the Milky Way. We do this assuming instant recycling, and that no r-process matter pollutes stars by any other means than these.

	In preparing these calculations we assumed a constant star formation rate, 2 M$_\odot yr^{-1}$, in the disk of the Milky Way, $\rm{r=15kpc}$, with a typical IMF and giant star birth rate of approximately $0.3\ $ stars per year. We also assumed the lifetime between $\log{g} = 3.2$ and the end of the core-helium burning phase is about $250\,{\rm Myr}$. Therefore if we assume a steady-state system, the number of giant stars in the Milky Way comes to approximately $7.53\times10^7$. This allows us to find the percentage of the Milky Ways stars that our sample represents.

	We also state that a star is said to be enhanced if [Eu/Fe] $\geq$ 0.7, and performed the lower-bound estimate that all of our 47 stars exhibit this abundance. The amount of r-process material expelled by neutron star mergers is the same as previously discussed, using Europium as our tracer, of $M_{Eu}\approx10^{-4.5} M_{\odot}$.

	These calculations were achieved by finding a density, $\rho$, such that:
	\begin{center}
	$\rho$ = $M_{Ejecta}*\rm{Rate}*13.8*10^{9}$
	\end{center}
	Where 'Rate' is defined as the number of neutron-star mergers that occur per parsec, per year.

	The total known mass of Europium exhibited in these 47 stars was then calculated:
	\begin{center}
	$Eu_{\rm{Mass}} = 47*Eu_{\rm{Atomic\ mass}}*N_{\rm{Eu atoms}}$\\\\
	\end{center}
	Where:
	\begin{center}
	$Eu_{\rm{Atomic\ mass}} = 2.52342^{-25}$ and;\\\\
	$N_{\rm{Eu\ atoms}} = 10^{[Eu/H]-12} \times N_{\rm{Hydrogen\ atoms}}$\\
	\end{center}
	Where the number of Hydrogen atoms was approximated from the Hydrogen atomic mass and the assumption that a giant star will be composed of $\approx$75\% Hydrogen and will have a mass of $\approx 0.8\ M_{\odot}$

    It was found that in the Milky Way there should be present 9,896 r-process stars enhanced solely from neutron star mergers. Taking into account that this should be considered a lower-bound estimate (by assuming a lower bound Europium abundance) this is a startlingly large difference from the known 61 r-process stars in the Milky Way.

    Scaling down from the 250 billion stars in the Milky Way to our sample size of $\approx450000$ giants, the amount of enhanced stars we expect to find comes to less than one (0.0178). If we are to extend our sample from the 450000 giants, and instead use the entire LAMOST data release 2 (a survey of $\approx2.2$ million stars), we predict that we should still have found less than a single star (0.087) enhanced in r-process elements. While again this should be considered a lower bound, it is nonetheless a startling difference from the 47 that were found.

    We can do a similar calculation with our rate of core-collapse supernovae. Taking the same assumptions with minimal mixing, as well as both the ejecta ($M_{Eu}\approx10^{-4.5} M_{\odot}$) and rate ($44700\ \rm{Gpc}^{-3} \rm{yr^{-1}}$) previously discussed, we find that in the Milky Way there should be present 442 r-process enhanced stars made solely from core-collapse supernovae.

    Were we to scale this down to our sample size of $\approx450000$ giants we arrive at a figure of $8.0\times10^{-4}$ stars that should have been observed. Scaling again to the full LAMOST sample we would then expect to find $3.9\times10^{-3}$ enhanced stars.

    From these estimates we can find an approximation for whether or not either neutron star mergers or core-collapse supernovae by themselves are sufficient to enrich our entire r-process enhanced catalogue.

    We find that, by using the assumptions listed, neutron star mergers could create enough r-process material to enrich above the point of [Eu/Fe] > 0.7 for the 47 known r-process enhanced stars within the LAMOST sample within $10.87\times10^9$ years, very close to the age of the universe. From our rates of core-collapse supernovae however, we find that the earliest time-scale at which they can create this level of r-process material is $2.4\times10^{11}$ years, not at all within the age of the universe. This illustrates, with our current knowledge of core-collapse supernovae frequency, and the amount of r-process material they eject, core-collapse supernovae cannot be the sole source of r-process enrichment in the Milky Way.

    This is reinforced by our r-process enhanced stars found in the LAMOST sample. Our 47 stars represent greater then 10\% of all enhanced stars present in the Milky Way if core-collapse supernovae is considered to be the singular source of r-process enhanced stars. However, the LAMOST data release 2 only surveyed 0.00088\% of the Milky Way. It is highly unlikely that 10\% of the entire r-process enhanced population of the Milky Way exists within our stellar sample, it is then concluded that core-collapse supernovae can not be the only source of r-process material.

    In an effort to confirm this, we approached the calculation from a different direction. Were we to take our sample of 47 giants found in the LAMOST data, and increase it to the full estimated amount of giants in the galaxy (75,300,000), then we would expect to find 7849 enhanced giant stars, very close to 9896 stars from the previous calculation.

	\section{Conclusions}
	We present the largest to date sample of r-process element enhanced stars to date. Our sample is found at a wide range of metallicities, contrary to previous discoveries that focussed on metal-poor r-process enhanced stars. These 47 stars represent the first r-process enhanced stars found at [Eu/Fe] $>-2$. They are not clumped in either temperatures, log(g), or radial velocity. They don't cluster to any one end of the H-R diagram, highlighting no correlation between type of star, they are observed both in the disk and halo, and are not clustered anywhere in the sky. In addition to this we also present the first ever estimate to our knowledge, from information provided by advanced LIGO, that neutron star mergers possess the ability to synthesise all currently known r-process material in a feasible time-scale. We also provide evidence that we can confidently conclude core-collapse supernovae can not possibly be the only progenitors of r-process material in the Milky Way Galaxy.

	We recommend follow-up high resolution spectra be obtained for each of the 47 stars to ascertain how enhanced they are. To find the origin of these stars and to show whether or not they may come from a similar progenitor site it is necessary to find orbits for them.

	% The best way to enter references is to use BibTeX:

	\bibliographystyle{mnras}
	%\bibliography{paper} % if your bibtex file is called example.bib
	\begin{thebibliography}{99}
		\bibitem[\protect\citeauthoryear{Abbott, B.P. et al.}{2016}]{LIGO2016}
		Abbot B.P. et al. 2016, Living Reviews in Relativity, 19, 1
		\bibitem[\protect\citeauthoryear{Argast, D. et al.}{2004}]{Argast2004}
		Argast D., Samland M., Thielmann F-K. \& Qian Y-Z. 2004, Astronomy and Astrophysics, 416
		\bibitem[\protect\citeauthoryear{Barklem, P.S. et al.}{2005}]{Barklem2005}
		Barklem, P.S., Christlieb, N., Beers, T.C., Hill, V., Besel, M.S., Holmberg, J., Marsteller, B., Rossi, S., Zickgraf, F.-J. \& Reimers, D. 2005, A\&A, 439, 1
		\bibitem[\protect\citeauthoryear{Burbidge, E. et al.}{1957}]{Burbidge1957}
		Burbidge E. M., Burbidge G. R., Fowler W. A., \& Hoyle F. 1957, Review of Modern Physics, 29, 4
		\bibitem[\protect\citeauthoryear{Burrows, A. et al.}{1995}]{Burrows1995}
		Burrows, A., Hayes, J., \& Fryxell, Bruce A. 1992, ApJ, 450, 830
		\bibitem[\protect\citeauthoryear{Cui, X.-Q. et al.}{2012}]{lamost}
		Cui, X.-Q. et al., Research in Astronomy and Astrophysics, 12, 9
		\bibitem[\protect\citeauthoryear{Drout, M. R. et al.}{2017}]{Drout2017}
		Drout, M. R. et al. 2017, Science
		\bibitem[\protect\citeauthoryear{Christlieb N. et al.}{2004}]{Christlieb2004}
		Christlieb N. et al. 2004, Astronomy \& Astrophysics, 428, 1
		\bibitem[\protect\citeauthoryear{Goriely, S. et al.}{2011}]{Goriely2011}
		Goriely S., Bauswein A. \& Janka H-T. 2011, , 738, L32
		\bibitem[\protect\citeauthoryear{Hayek, W. et al.}{2009}]{Hayek2009}
		Hayek W., Wiesendahl U., Christlieb N., Eriksson K., Barklem P. S., Hill V., Beers T. C., Farouqi K., Pfeiffer B.,\& Kratz K.-L. 2009, Astronomy and Astrophysics, 504, 1
		\bibitem[\protect\citeauthoryear{Hill, V. et al.}{2002}]{Hill2002}
		Hill V. et al. Astronomy \& Astrophysics, 387, 1
		\bibitem[\protect\citeauthoryear{Ho et al.}{2017}]{AnnaHo2017}
		Ho A. Y. Q., Ness M. K., Hogg D. W., Rix H-W., Liu C., Yang F., Zhang Y., Hou Y.,\& Wang Y. 2017, ApJ, 836, 1
		\bibitem[\protect\citeauthoryear{Honda S. et al.}{2004}]{Honda2004}
		Honda S., Aoki W., Kajino T., Ando H., Beers T. C., Izumiura H., Sadakane K. \& Takada-Hidai M. 2004, ApJ, 607, 1
		\bibitem[\protect\citeauthoryear{Hotokezaka, K. et al.}{2013}]{Hotok2013}
		Hotokezaka K., Kiuchi K., Kyutoku K., Okawa H., Skiguchi Y., Shibata M., \& Taniguchi K. 2013, APS
		\bibitem[\protect\citeauthoryear{Ivans, I. I. et al.}{2006}]{Ivans2006}
		Ivans I. I.,Simmerer J., Sneden C., Lawler J. E., Cowan J. J., Gallino R., \& Bisterzo S. 2006, ApJ, 645, 1
		\bibitem[\protect\citeauthoryear{Ji, A. P. et al.}{2016}]{Ji2016}
		Ji A. P., Frebel A., Chiti A., \& Simon J. D. 2016, Nature, 531
		\bibitem[\protect\citeauthoryear{Kasen, D. et al.}{2017}]{Kasen2017}
		Kasen D., Metzger B., Barnes J., Quataert E., \& Ramirez-Ruiz E. 2017, Nature, 551, 7678
		\bibitem[\protect\citeauthoryear{Lai, D. K. et al.}{2008}]{Lai2008}
		Lai D. K., Bolte M., Johnson J. A., Lucatello S., Heger A., \& Woosley S. E. 2008, ApJ, 681,
		\bibitem[\protect\citeauthoryear{Li, W. et al.}{2011}]{Li2011}
		Li W., Chornock R., Leaman J., Filippenko A. V., Poznanski D., Wang X., Ganshalingam M., \& Mannucci F. 2011, Royal Astronomical Society, 412, 1473
		\bibitem[\protect\citeauthoryear{Majewski, S.R., et al.}{2017}]{apogee}
		Majewski S.R., et al., The Astronomical Journal, 154, 3
		\bibitem[\protect\citeauthoryear{Mashonkina, L., et al.}{2010}]{masho2011}
		Mashonkina L., Christlieb N., Barklem P.S., Hill V., Beers T.C., \& Velichko A., Astronomy and Astrophysics, 516, 46
		\bibitem[\protect\citeauthoryear{Placco, V.M., et al.}{2017}]{placo2017}
		Placco V.M., Holmbeck E. M., Frebel A., Beers T. C., Surman R. A., Ji A. P., Ezzeddine R., Points S. D., Kaleida C. C. Hansen T. T., Sakari C. M, \& Casey A. R., ApJ, 844, 1
		\bibitem[\protect\citeauthoryear{Sneden, C. et al.}{2008}]{Sneden2008}
		Sneden C., Cowan J.J., \& Gallino R. 2008, Annual review of Astronomy \& Astrophysics, 46, 1
		\bibitem[\protect\citeauthoryear{Sneden, C. et al.}{2003}]{Sneden2003}
		Sneden C. et al. 2003, ApJ, 591, 2
		\bibitem[\protect\citeauthoryear{Wako, A. et al.}{2010}]{Wako2010}
		Wako A., Beers T. C., Honda S. \& Carollo D. 2010, ApJ, 723, 1
		\bibitem[\protect\citeauthoryear{Woosley, S.E. et al.}{1992}]{Woosley1992}
		Woosley S.E, \& Hoffman R.D. 1992, ApJ, 395, 1
	\end{thebibliography}

	% Please add the following required packages to your document preamble:
	% \usepackage{booktabs}
	% Please add the following required packages to your document preamble:
	% \usepackage{booktabs}
	\begin{table*}
		\centering
		\label{Data for the 47}
		\begin{tabular}{@{}cccccccccccc@{}}
			\toprule
			\textbf{2MASS ID}   & \textbf{RA} & \textbf{DEC} & \textbf{S/N}        	& \textbf{$\boldsymbol{V_{\rm{r}}}$}   & \textbf{$\boldsymbol{T_{\rm{eff}}}$} & \textbf{Log(g)}	& \textbf{$\boldsymbol{[\rm{Fe/H}]}$} & \textbf{$\boldsymbol{[\alpha/\rm{H}]}$} & \textbf{$\boldsymbol{\chi_{\rm{r}}^{2}}$} & \textbf{{[}Eu/Fe{]}} & \textbf{Error} \\
			-               	& {[}H:M:S{]}   & {[}H:M:S{]}	& {[}$\rm{pixel}^{-1}${]} & {[}$\rm{km\ s}^{-1}${]} & {[}K{]}             	& {[}$\rm{cm\ s}^{-2}${]} & {[}dex{]}          	& {[}dex{]}              	& -                        	& {[}dex{]}        	& {[}dex{]}  	\\ \midrule
			J130303.65+260837.0 & 13:03:03.65 & +26:08:37.0  & 36.48               	& -37.77              	& 4851.24             	& 2.98           	& -0.54              	& 0.19                   	& 0.42                     	& 1.39             	& 0.31       	\\
			J062733.76+280814.1 & 06:27:33.77 & +28:08:14.2  & 39.08               	& 38.67               	& 4502.58             	& 1.92           	& -0.15              	& 0.00                   	& 0.35                     	& 1.43             	& 0.35       	\\
			J070035.19+103443.5 & 07:00:35.19 & +10:34:43.5  & 41.77               	& 17.69               	& 4752.08             	& 2.44           	& -0.43              	& 0.06                   	& 0.33                     	& 1.09             	& 0.63       	\\
			J070107.18+125948.2 & 07:01:07.19 & +12:59:48.2  & 33.72               	& 46.17               	& 4452.61             	& 2.51           	& 0.47               	& 0.02                   	& 0.36                     	& 1.30             	& 0.31       	\\
			J063400.95+420904.3 & 06:34:00.96 & +42:09:04.4  & 34.09               	& 9.89                	& 4315.56             	& 1.62           	& -0.50              	& 0.10                   	& 0.43                     	& 1.19             	& 3.10       	\\
			J220118.66+033558.6 & 22:01:18.66 & +03:35:58.6  & 35.11               	& -108.23             	& 4790.06             	& 2.51           	& -1.36              	& 0.33                   	& 0.37                     	& 1.44             	& 0.26       	\\
			J221220.27+032735.4 & 22:12:20.27 & +03:27:35.5  & 35.16               	& -12.89              	& 4867.34             	& 3.12           	& 0.11               	& 0.14                   	& 0.35                     	& 1.52             	& 0.20       	\\
			J060021.29+384840.6 & 06:00:21.29 & +38:48:40.6  & 30.03               	& 32.68               	& 4305.08             	& 2.10           	& 0.15               	& -0.01                  	& 0.54                     	& 1.49             	& 0.35       	\\
			J081632.10-053703.4 & 08:16:32.11 & -05:37:03.4  & 59.16               	& 44.67               	& 4541.14             	& 2.45           	& 0.57               	& 0.03                   	& 0.75                     	& 1.34             	& 0.26       	\\
			J064641.14+235318.2 & 06:46:41.14 & +23:53:18.3  & 36.77               	& 25.78               	& 4325.56             	& 1.78           	& -0.10              	& 0.02                   	& 0.42                     	& 1.41             	& 0.27       	\\
			J025225.88+324308.1 & 02:52:25.88 & +32:43:08.2  & 34.79               	& 24.88               	& 4797.58             	& 2.52           	& -0.23              	& 0.05                   	& 0.24                     	& 1.31             	& 0.29       	\\
			J061822.02+015126.7 & 06:18:22.02 & +01:51:26.8  & 36.75               	& 45.27               	& 4617.95             	& 2.43           	& -0.12              	& 0.04                   	& 0.34                     	& 1.42             	& 0.30       	\\
			J083049.21+304144.3 & 08:30:49.21 & +30:41:44.3  & 39.5                	& -125.31             	& 4859.92             	& 2.28           	& -1.63              	& 0.33                   	& 0.41                     	& 1.52             	& 0.20       	\\
			J035949.12+302104.4 & 03:59:49.13 & +30:21:04.4  & 30.42               	& 26.98               	& 4164.41             	& 1.79           	& -0.12              	& 0.00                   	& 0.36                     	& 1.47             	& 0.32       	\\
			J074852.24+075016.5 & 07:48:52.25 & +07:50:16.6  & 33.47               	& 36.87               	& 4405.84             	& 1.94           	& -0.48              	& 0.07                   	& 0.30                     	& 1.44             	& 0.30       	\\
			J163116.81+002349.5 & 16:31:16.81 & +00:23:49.5  & 32.63               	& -35.98              	& 4628.92             	& 2.53           	& 0.01               	& 0.18                   	& 0.29                     	& 1.44             	& 0.32       	\\
			J073532.54+215850.0 & 07:35:32.55 & +21:58:50.0  & 35.61               	& 35.98               	& 4859.66             	& 3.27           	& 0.19               	& 0.06                   	& 0.44                     	& 1.19             	& 0.79       	\\
			J164023.01+164552.5 & 16:40:23.02 & +16:45:52.6  & 44.59               	& -34.18              	& 4964.91             	& 2.96           	& 0.36               	& 0.07                   	& 0.35                     	& 1.27             	& 0.32       	\\
			J153333.12+510621.8 & 15:33:33.13 & +51:06:21.9  & 49.45               	& -17.99              	& 4861.78             	& 2.51           	& -0.49              	& 0.22                   	& 0.46                     	& 1.31             	& 0.28       	\\
			J175737.71+521523.1 & 17:57:37.71 & +52:15:23.2  & 38.52               	& -40.47              	& 4861.38             	& 3.36           	& -0.34              	& 0.10                   	& 0.55                     	& 1.51             	& 0.24       	\\
			J184125.86+432807.8 & 18:41:25.86 & +43:28:07.9  & 46.94               	& -93.84              	& 4703.92             	& 2.61           	& -0.26              	& 0.27                   	& 0.33                     	& 1.20             	& 0.34       	\\
			J195234.41+470859.4 & 19:52:34.41 & +47:08:59.5  & 68.57               	& -24.28              	& 4596.44             	& 2.54           	& 0.55               	& 0.07                   	& 0.82                     	& 0.95             	& 0.48       	\\
			J215146.68+304016.0 & 21:51:46.68 & +30:40:16.0  & 73.25               	& -28.78              	& 4566.21             	& 2.48           	& 0.64               	& 0.04                   	& 0.71                     	& 0.88             	& 0.45       	\\
			J075116.55+220137.8 & 07:51:16.56 & +22:01:37.9  & 32.87               	& 90.24               	& 4810.60             	& 2.70           	& -0.33              	& 0.02                   	& 0.30                     	& 1.35             	& 0.31       	\\
			J003929.43+430039.5 & 00:39:29.43 & +43:00:39.6  & 33.14               	& -104.33             	& 4190.26             	& 1.87           	& -0.01              	& 0.07                   	& 0.41                     	& 1.35             	& 0.33       	\\
			J040122.07+461504.5 & 04:01:22.08 & +46:15:04.5  & 35.43               	& -95.03              	& 4305.88             	& 1.75           	& 0.02               	& 0.10                   	& 0.57                     	& 1.37             	& 0.35       	\\
			J040815.45+464920.1 & 04:08:15.46 & +46:49:20.1  & 33.05               	& 16.19               	& 4855.16             	& 2.65           	& -0.11              	& 0.00                   	& 0.30                     	& 1.48             	& 0.30       	\\
			J071206.83+213038.0 & 07:12:06.84 & +21:30:38.1  & 32.66               	& 29.08               	& 4432.03             	& 1.75           	& -0.38              	& 0.11                   	& 0.40                     	& 1.47             	& 0.28       	\\
			J222820.22+340750.3 & 22:28:20.22 & +34:07:50.3  & 34.52               	& -45.27              	& 4419.96             	& 2.06           	& -0.07              	& 0.11                   	& 0.49                     	& 1.37             	& 0.31       	\\
			J222310.49+344357.4 & 22:23:10.50 & +34:43:57.5  & 54.06               	& -20.09              	& 4232.66             	& 1.87           	& -0.01              	& 0.03                   	& 0.86                     	& 1.22             	& 0.31       	\\
			J222101.61+365923.6 & 22:21:01.62 & +36:59:23.7  & 30.81               	& -27.88              	& 4360.28             	& 2.23           	& 0.20               	& 0.04                   	& 0.45                     	& 1.47             	& 0.29       	\\
			J052327.44+571126.8 & 05:23:27.45 & +57:11:26.8  & 33.54               	& -24.88              	& 4806.24             	& 3.08           	& 0.24               	& 0.07                   	& 0.45                     	& 1.43             	& 0.25       	\\
			J235212.27+492548.4 & 23:52:12.27 & +49:25:48.4  & 39.93               	& -95.03              	& 5047.36             	& 3.30           	& -0.43              	& 0.22                   	& 0.29                     	& 1.28             	& 0.26       	\\
			J074102.63+130520.3 & 07:41:02.64 & +13:05:20.4  & 40.83               	& 49.17               	& 5000.84             	& 3.00           	& -0.38              	& -0.02                  	& 0.30                     	& 1.35             	& 0.27       	\\
			J061559.10+470344.9 & 06:15:59.11 & +47:03:44.9  & 36.84               	& -18.29              	& 4811.98             	& 2.82           	& -0.59              	& 0.07                   	& 0.44                     	& 1.00             	& 1.33       	\\
			J054704.12+600827.1 & 05:47:04.13 & +60:08:27.1  & 33.19               	& -23.33              	& 4770.80             	& 2.64           	& -0.10              	& 0.06                   	& 0.40                     	& 1.46             	& 0.26       	\\
			J055548.33+611537.1 & 05:55:48.33 & +61:15:37.1  & 33.6                	& -40.88              	& 4850.39             	& 2.57           	& -0.46              	& 0.13                   	& 0.22                     	& 1.42             	& 0.26       	\\
			J042723.73+271546.8 & 04:27:23.73 & +27:15:46.9  & 45.73               	& -40.17              	& 4211.91             	& 1.72           	& -0.04              	& 0.05                   	& 0.79                     	& 1.33             	& 0.34       	\\
			J035702.05+281335.2 & 03:57:02.06 & +28:13:35.3  & 31.82               	& -44.67              	& 4712.89             	& 2.44           	& 0.01               	& 0.16                   	& 0.46                     	& 1.28             	& 0.38       	\\
			J031030.13+531759.9 & 03:10:30.13 & +53:17:59.9  & 35.34               	& -16.79              	& 4699.75             	& 2.34           	& 0.21               	& 0.00                   	& 0.89                     	& 1.51             	& 0.21       	\\
			J091256.32+353008.3 & 09:12:56.32 & +35:30:08.3  & 73.39               	& 20.39               	& 4441.95             	& 2.33           	& 0.41               	& 0.02                   	& 0.76                     	& 0.92             	& 0.39       	\\
			J071534.00+211353.8 & 07:15:34.00 & +21:13:53.9  & 31.68               	& 42.27               	& 4892.00             	& 2.57           	& -0.72              	& 0.11                   	& 0.34                     	& 0.59             	& 0.29       	\\
			J101317.14+122928.8 & 10:13:17.14 & +12:29:28.8  & 45.54               	& 300.69              	& 5055.95             	& 2.24           	& -1.34              	& 0.33                   	& 0.38                     	& 1.46             	& 0.24       	\\
			J161326.00+073218.1 & 16:13:26.00 & +07:32:18.2  & 48.45               	& -50.07              	& 5086.79             	& 3.40           	& -0.68              	& 0.29                   	& 0.93                     	& 1.38             	& 0.24       	\\
			J173711.85+101126.6 & 17:37:11.85 & +10:11:26.7  & 30.01               	& -33.58              	& 4991.07             	& 3.49           	& -0.05              	& 0.17                   	& 0.41                     	& 1.45             	& 0.28       	\\
			J173041.92+125640.5 & 17:30:41.92 & +12:56:40.5  & 52.67               	& 20.39               	& 4986.98             	& 3.42           	& -0.45              	& 0.22                   	& 0.30                     	& 1.09             	& 0.43       	\\
			J172929.42+120133.3 & 17:29:29.43 & +12:01:33.3  & 39.91               	& 14.39               	& 4597.08             	& 2.67           	& 0.24               	& 0.05                   	& 0.67                     	& 1.02             	& 0.80       	\\ \bottomrule
		\end{tabular}
		\caption{Abundances and additional information of the 47 enhanced stars.}
	\end{table*}

	% Don't change these lines
	%\bsp	% typesetting comment
	\label{lastpage}
\end{document}